A Mountaintop View Requires Minimal Sorting: \break A Faster Contour Tree Algorithm \footnote{The full updated version of this paper is available online \url{\si{http://web.engr.illinois.edu/\string~\si{raichel2}/\si{paint.pdf}}}}

### 2014-11-04 paint-contour-tree.tex
sentences added

@@

We complement \Thm{main-alg} with a lower bound, suggesting that the comparison complexity
of $\sum_{p \in P(T)} |p|\log |p|$ is inherent for any contour tree algorithm. The exact phrasing
of the lower bound requires some care. Consider the $d=2$ (terrain) case.
%
%
We define a class of valid path decompositions.  This include any decomposition of rooted trees,
and more generally is a large superset of all decomposition produced by our 
algorithm.
%\footnote{Our construction likely works for any decomposition, however, 
%to keep the presentation simple we restrict to a subset relevant for our algorithm.}
%
%
%Let a path decomposition $P$ be called \emph{valid} if it is actually output by the algorithm
%on some input.
%
%
%As referenced in \Thm{main-alg}, for any $f:\MM \mapsto \RR$ our deterministic algorithm defines a 
%specific path decomposition $P(T)$%
\ignore{\FullVer{
\footnote{The specific decomposition used is defined in \Sec{pathDecomp}, 
but also depends on the raining and painting procedures described in \Sec{rain} and \Sec{paint}.  
The important thing is that our algorithm is deterministic, and so for a fixed $\MM$ and $f$ it is well defined.}.}
\SubmitVer{
\footnote{The specific decomposition used is defined in the full version of the paper.  
The important thing is that our algorithm is deterministic, and so for a fixed $\MM$ and $f$ it is well defined.}.}
We define $\cost(f) = \sum_{p \in P(T)} |p|\log |p|$ for this specific decomposition $P(T)$.
Ideally, we would like to argue that any contour-tree algorithm
must pay $\cost(f)$ for ant $f$. 
But this is obviously false, since one could consider an algorithm
that hard codes the answer for $f$ and merely checks this answer.

>>>>

In particular, any comparisons made are only between ancestors and descendants in the contour tree, where as previous algorithm compared everything, 
and that is where we gain.

@@

We complement \Thm{main-alg} with a lower bound, suggesting that the comparison complexity
of $\sum_{p \in P(T)} |p|\log |p|$ is inherent for any contour tree algorithm. The exact phrasing
of the lower bound requires some care. Consider the $d=2$ (terrain) case.
%
We define a class of valid path decompositions which characterizes decompositions produced by our algorithm.
%
%We define a class of valid path decompositions.  This include any decomposition of rooted trees,
%and more generally is a large superset of all decomposition produced by our 
%algorithm.
%\footnote{Our construction likely works for any decomposition, however, 
%to keep the presentation simple we restrict to a subset relevant for our algorithm.}
%
%Let a path decomposition $P$ be called \emph{valid} if it is actually output by the algorithm
%on some input.
%
%
%As referenced in \Thm{main-alg}, for any $f:\MM \mapsto \RR$ our deterministic algorithm defines a 
%specific path decomposition $P(T)$%
\ignore{\FullVer{
\footnote{The specific decomposition used is defined in \Sec{pathDecomp}, 
but also depends on the raining and painting procedures described in \Sec{rain} and \Sec{paint}.  
The important thing is that our algorithm is deterministic, and so for a fixed $\MM$ and $f$ it is well defined.}.}
\SubmitVer{
\footnote{The specific decomposition used is defined in the full version of the paper.  
The important thing is that our algorithm is deterministic, and so for a fixed $\MM$ and $f$ it is well defined.}.}
We define $\cost(f) = \sum_{p \in P(T)} |p|\log |p|$ for this specific decomposition $P(T)$.
Ideally, we would like to argue that any contour-tree algorithm
must pay $\cost(f)$ for ant $f$. 
But this is obviously false, since one could consider an algorithm
that hard codes the answer for $f$ and merely checks this answer.
